---
title: "Analysis for Mindfulness-based Cognitive Training Study"
author: "Chloe Dziego"
date: "27/02/2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Data tidying/visualising packages
library(tidyverse)
library(moments)
library(rstatix)
library(dplyr)
library(ggpubr)
library(skimr)
library(psych)
library(eegUtils)

#Packages for statistics
library(emmeans)
library(ggeffects)
library(performance)
library(lme4)
library(splines)
library(car)
library(lmerOut)
library(lmerTest)

options(contrasts = c("contr.Sum","contr.helmert"))
options(decorate.contr.Treatment="",decorate.contr.Sum="",decorate.contrasts=c("[", "]"))

```

```{r raincloud plot function}
# Defining the geom_flat_violin function ----

# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50

"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}
geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}
#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          draw_key = draw_key_polygon,
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          required_aes = c("x", "y")
  )

outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          cat("Nothing changed", "n")
          return(invisible(var_name))
     }
}
```

```{r read in data + tidy}

##Prepare Neural and Performance Variable Data
iaf <- read.delim("iaf_output.txt", header = TRUE)
iaf$subj <- as.factor(iaf$subj)
iaf$session <- as.factor(iaf$session)
iaf <- pivot_wider(iaf, names_from = measure, values_from = value)

aperiodic <- read.csv("aperiodic.csv") %>%
  dplyr::select(Chan, Intercept, Slope, subj, session) %>%
  filter(Chan == "Pz" | Chan == "P3" | Chan == "P4" | Chan == "O1" | Chan == "Oz" | Chan == "O2" | Chan == "P7" | Chan == "P8")

mean_sl <- aggregate(aperiodic$Slope, by = list(aperiodic$subj, aperiodic$session), FUN = mean) %>%
  dplyr::rename("subj" = Group.1,
         "session" = Group.2,
         "slope" = x)

mean_int <- aggregate(aperiodic$Intercept, by = list(aperiodic$subj, aperiodic$session), FUN = mean) %>%
  dplyr::rename("subj" = Group.1,
         "session" = Group.2,
         "intercept" = x)

tpe <- read_csv("all_tpe.csv") %>%
  na.omit()

tpe$session <- as.factor(tpe$session)
tpe$subj <- as.factor(tpe$subj)

all_ap <- left_join(mean_sl, mean_int, by = c("subj", "session"))
all_ap$session <- as.factor(all_ap$session)

iaf_ap <- left_join(all_ap, iaf, by = c("subj", "session"))
almost <- left_join(tpe, iaf_ap, by = c("subj", "session"))
  
data_woct <- subset(almost, Time!="0") %>%
  filter(subj != "CT22" & subj != "CT25")

data_woct$subj <- as.factor(data_woct$subj)

check <- data_woct %>%
  group_by(subj) %>%
  count() 

```

``` {r code topoplots}

#Create Topoplots for Visualising Neural Variables across the Scalp
library(eegUtils)

for_topo1 <- read.csv("aperiodic.csv") %>%
  dplyr::select(Chan, Intercept, Slope, subj, session) %>%
  filter(session == 1) %>%
  rename(electrode = Chan) %>%
  group_by(electrode) %>%
  summarise(amplitude = mean(Slope))

for_topo2 <- read.csv("aperiodic.csv") %>%
  dplyr::select(Chan, Intercept, Slope, subj, session) %>%
  filter(session == 2) %>%
  rename(electrode = Chan) %>%
  group_by(electrode) %>%
  summarise(amplitude = mean(Slope))

for_topo1$electrode <- as.factor(for_topo1$electrode)

topoplot(for_topo1, palette = "GnBu", limits = c(-1.5, -0.25))
topoplot(for_topo2, palette = "GnBu", limits = c(-1.5, -0.25))

int_topo1 <- read.csv("aperiodic.csv") %>%
  dplyr::select(Chan, Intercept, Slope, subj, session) %>%
  filter(session == 1) %>%
  rename(electrode = Chan) %>%
  group_by(electrode, session) %>%
  summarise(amplitude = mean(Intercept))

int_topo2 <- read.csv("aperiodic.csv") %>%
  dplyr::select(Chan, Intercept, Slope, subj, session) %>%
  filter(session == 2) %>%
  rename(electrode = Chan) %>%
  group_by(electrode) %>%
  summarise(amplitude = mean(Intercept))

int_topo1$electrode <- as.factor(int_topo1$electrode)

topoplot(int_topo1, palette = "GnBu", limits = c(-30, -20))
topoplot(int_topo2, palette = "GnBu", limits = c(-30, -20))

recodedf <- structure(list(V1 = c('Fz',
 'F3',
 'F7',
 'FT9',
 'FC5',
 'FC1',
 'C3',
 'T7',
 #'TP9', is number 8
 'CP5',
 'CP1',
 'Pz',
 'P3',
 'P7',
 'O1',
 'Oz',
 'O2',
 'P4',
 'P8',
 #'TP10', is number 19
 'CP6',
 'CP2',
 'Cz',
 'C4',
 'T8',
 'FT10',
 'FC6',
 'FC2',
 'F4',
 'F8'), V2 = c(0:8, 9:18, 20:28)), .Names = c("V1", 
    "V2"), class = "data.frame", row.names = c(NA, -29L))

recodedf <- aperiodic %>%
  select(X, Chan) %>%
  rename(V2 = X, V1 = Chan) %>%
  distinct()

recodedf$V2 <- as.factor(recodedf$V2)

#IAF per electrode
iafelectrode <- read.delim("iaf_electrodes.txt", header = TRUE) %>%
  rename(channel = electrode)

iafelectrode$value <- as.numeric(iafelectrode$value)
iafelectrode$electrode <- recodedf$V1[match(as.character(iafelectrode$channel), as.character(recodedf$V2))]
iafelectrode$electrode <- as.factor(iafelectrode$electrode)

iaf_topo1 <- iafelectrode %>%
  dplyr::select(subj, measure, value, session, electrode) %>%
  filter(measure == "cog") %>%
  filter(session == 1) %>%
  group_by(electrode) %>%
  summarise(across(c('value'), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}")) %>%
  rename(amplitude = mean_value)

iaf_topo2 <- iafelectrode %>%
  dplyr::select(subj, measure, value, session, electrode) %>%
  filter(measure == "cog") %>%
  filter(session == 2) %>%
  group_by(electrode) %>%
  summarise(across(c('value'), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}")) %>%
  rename(amplitude = mean_value)

topoplot(iaf_topo1, palette = "GnBu", limits = c(8, 11))
topoplot(iaf_topo2, palette = "GnBu", limits = c(8, 11))

iaf_topo12 <- iafelectrode %>%
  dplyr::select(subj, measure, value, session, electrode) %>%
  filter(measure == "paf") %>%
  filter(session == 1) %>%
  group_by(electrode) %>%
  summarise(across(c('value'), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}")) %>%
  rename(amplitude = mean_value)

iaf_topo22 <- iafelectrode %>%
  dplyr::select(subj, measure, value, session, electrode) %>%
  filter(measure == "paf") %>%
  filter(session == 2) %>%
  group_by(electrode) %>%
  summarise(across(c('value'), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}")) %>%
  rename(amplitude = mean_value)

topoplot(iaf_topo12, palette = "GnBu", limits = c(8, 11))
topoplot(iaf_topo22, palette = "GnBu", limits = c(8, 11))

```

## Report Descriptives & Plots

```{r descriptives}

#Create Descriptive Tables
neural1 <- filter(iaf_ap, session == "1")
neural2 <- filter(iaf_ap, session == "2")

describe(neural1)
describe(neural2)

meanslopes <- aggregate(iaf_ap$slope, by = list(iaf_ap$session), FUN = mean)%>%
  dplyr::rename("session" = Group.1,
         "slope" = x)
meanint <- aggregate(iaf_ap$intercept, by = list(iaf_ap$session), FUN = mean)%>%
  dplyr::rename("session" = Group.1,
         "intercept" = x)

aperiodicmeans <- left_join(meanslopes, meanint, by = "session")

#Descriptives for Demographic Information
demographics_all <- read_csv("demo.csv")

demographics_all$sex <- as.factor(demographics_all$sex)
demographics_all$scenario <- as.factor(demographics_all$scenario)
summary(demographics_all)
skim(demographics_all)

#These subjects did not complete two sessions
demographics <- demographics_all %>% filter(subj != "CT22" & subj != "CT25")

summary(demographics)
scenario_file <- demographics %>%
  select(subj, scenario)

```

## Cognitive Training Adherence
Number of times participant engaged in the cognitive training + MAAS. 

```{r ct adherence}

ct <- read_csv("ct_tracker.csv") %>%
  select(subj, total)

ct_all <- read_csv("ct_tracker.csv") %>%
  select(subj, day1, day2, day3, day4, day5, day6, day7) %>%
  filter(subj != "CT22" & subj != "CT25")

day <- c(1:7)
mean <- c(1.15, 1.5, 1.45, 1.5, 1.625, 1.6, 1.45)
daily_avg <- data.frame(day, mean)

summary(daily_avg)
ggplot(daily_avg, aes(x = day, y = mean)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(lim = c(0, 3)) +
  scale_x_continuous(n.breaks = 7) +
  labs(x = "Day of Training", y = "Average Sessions Completed") +
  theme_minimal() +
  stat_regline_equation()

summary(lm(mean ~ day, data = daily_avg))

ct_mSD <- ct_all %>%
  rowwise() %>%
  mutate(mean = mean(c_across(day1:day7)),
         SD = SD(c_across(day1:day7)))
  
skews = c()

for(i in 1:nrow(ct_mSD)) {
    row <- ct_mSD[i,]
    sum1 <- (((row$day1)-(row$mean))/row$SD)^3
    sum2 <- (((row$day2)-(row$mean))/row$SD)^3
    sum3 <- (((row$day3)-(row$mean))/row$SD)^3
    sum4 <- (((row$day4)-(row$mean))/row$SD)^3
    sum5 <- (((row$day5)-(row$mean))/row$SD)^3
    sum6 <- (((row$day6)-(row$mean))/row$SD)^3
    sum7 <- (((row$day7)-(row$mean))/row$SD)^3
    sum <- sum(sum1, sum2, sum3, sum4, sum5, sum6, sum7) 
    skew_value <- (1/7)*sum
    
    skews[i] <- skew_value[1] 
}

ct_mSD[11] <- skews
ct_skew <- ct_mSD %>%
  rename(skew = ...11)

maas_all <- read_csv("maas_scores.csv")

#Remove Participants who were pilot/did not end up participating/or completing the full two sessions
maas <- maas_all[c(3:52),18:30] %>%
  filter(Q1 != "EXP99" & Q1 != "EXP02" & Q1 != "EXP01" & Q1 != "CT22" & Q1 != "CT25" & Q1 != "CT14"& Q1 != "EXP03" & Q1 != "EXP05" ) %>%
  na.omit() %>%
  mutate_at(vars(2:ncol(.)), list(~
  case_when(
    . == "Almost Always" ~ 1,
    . == "Very Frequently" ~ 2,
    . == "Somewhat Frequently" ~ 3,
    . == "Somewhat Infrequently" ~ 4,
    . == "Very Infrequently" ~ 5,
    . == "Almost Never" ~ 6))) %>%
  dplyr::rename("subj"=Q1)

maas$MAAS <- rowSums(maas[2:12])/12
maas.2 <- select(maas, subj, MAAS)

summary(maas.2$MAAS)

alpha(maas[2:13])
describe(maas.2)
ggplot(maas.2, aes(x=MAAS)) +
  geom_histogram()

describe(ct)

data <- left_join(data_woct, ct, by = c("subj")) %>%
  dplyr::rename("training" = total)
data$TPE <- as.integer(data$TPE)

objective_report <- ggplot(ct, aes(subj, total)) + 
  geom_col(fill = c("seagreen3"), alpha = 0.4) +
  ylab("Number of Sessions") +
  scale_y_continuous(limits = c(0, 20)) +
  ggtitle("A. Engagement (Analytics)") +
  theme_minimal() +
  theme(axis.title.y= element_text(size = 16),
        axis.title.x= element_blank(),
        axis.text.y= element_text(size = 15),
        axis.text.x= element_text(size = 14),
        plot.title = element_text(size=16, face="bold"))

tiff("frequency_of_engagement.tiff", units="in", width=10, height=7, res=300)

ggplot(ct, aes(x=total)) +
  geom_histogram(fill = "turquoise2", alpha = 0.4) +
  labs(title= "Frequency Histogram of Cognitive Training Adherence", x = "Audio Tracks Completed", y = "Number of Participants") +
  theme_minimal() +
  theme(axis.text.x=element_text(size = "16"),
        axis.title.x = element_text(size ="16"),
        axis.text.y =element_text(size = "16"),
        axis.title.y = element_text(size ="16"))
dev.off()

#New Data Frame with Low/High Dichotomies
ct_extra <- mutate(ct, level = case_when(total > 10 ~ "High",
                                   total < 10 ~ "Low")) %>%
   filter(subj != "CT22" & subj != "CT25")

#Create dataframe of neural variables and cognitive training data
iaf_ct <- left_join(iaf, ct_extra, by = "subj") %>%
  na.omit()


```

## Initial Data Exploration

First visualisation of neural metrics change across session 1 and 2.

```{r data exploration neural}

paf.plot <- ggplot(iaf, aes(x = session, y = paf, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = 0.3) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(group = subj), size = 2, shape = 19, alpha = 0.2, colour = "black", position = position_dodge(0.4)) +
  ylab("Individual Alpha Frequency (IAF)") +
  xlab("Session") +
  ggtitle("C. Peak Alpha Frequency") +
  scale_x_discrete(labels=c("Pre-Meditiation", "Post-Meditation")) +
  theme_bw() + 
  theme(legend.position="none",
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.text.x = element_text(size = 14),
        plot.title = element_text(size=14, face="bold"),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_blank())

paf.plot

cog.plot <- ggplot(iaf, aes(x = session, y = cog, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = 0.3) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(fill = session, group = subj), size = 2, shape = 19, alpha = 0.2, colour = "black", position = position_dodge(0.4)) +
  ylab("Individual Alpha Frequency (IAF)") +
  xlab("Session") +
  scale_x_discrete(labels=c("Pre-Meditiation", "Post-Meditation")) +
  scale_y_continuous(limits = c(9.5, 12)) +
  ggtitle("D. Centre of Gravity") +
  theme_bw() + 
  theme(legend.position="none",
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size=14, face="bold"),
        axis.title.x = element_blank())

tiff("iaf_session.tiff", units="in", width=12, height=6, res=300)
ggarrange(paf.plot, cog.plot, common.legend = FALSE)
dev.off()

mean_sl$session <- as.factor(mean_sl$session)
mean_int$session <- as.factor(mean_int$session)

slope.plot <- ggplot(mean_sl, aes(x = session, y = slope, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = 0.3) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(group = subj), size = 2, shape = 19, alpha = 0.2, colour = "black", position = position_dodge(0.4)) +
  ylab("Aperiodic 1/f Slope") +
  ggtitle("E. 1/f Slope") +
  scale_x_discrete(labels=c("Pre-Meditiation", "Post-Meditation")) +
  theme_bw() + 
  theme(legend.position="none",
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size=14, face="bold"),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_blank())

slope.plot

intercept.plot <- ggplot(mean_int, aes(x = session, y = intercept, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = 0.3) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(fill = session, group = subj), size = 2, shape = 19, alpha = 0.2, colour = "black", position = position_dodge(0.4)) +
  ylab("Aperiodic 1/f Intercept") +
  ggtitle("F. 1/f Intercept") +
  scale_x_discrete(labels=c("Pre-Meditiation", "Post-Meditation")) +
  theme_bw() + 
  theme(legend.position="none",
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.text.x = element_text(size = 14),
        plot.title = element_text(size=14, face="bold"),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_blank())

intercept.plot

tiff("aperiodic_session.tiff", units="in", width=12, height=6, res=300)
ggarrange(slope.plot, intercept.plot, common.legend = FALSE)
dev.off()

#t-tests of neural metrics from time 1 to 2
ttest_slope <- pivot_wider(iaf_ap, id_cols = subj, names_from = session, values_from = slope)
ttest_intercept <- pivot_wider(iaf_ap, id_cols = subj, names_from = session, values_from = intercept)
ttest_cog <- pivot_wider(iaf_ap, id_cols = subj, names_from = session, values_from = cog)

t.test(ttest_slope$'1', ttest_slope$'2', paired = TRUE)
t.test(ttest_intercept$'1', ttest_intercept$'2', paired = TRUE)
t.test(ttest_cog$'1', ttest_cog$'2', paired = TRUE)

```

## Visualisation of TPE 

Examining changes in TPE across session 1 and 2.

```{r tpe visualisation}

group_labels = c("Pre-Meditation", "Post-Meditation")

#All participant individual TPE scores over time
ggplot(data, aes(x = Time, y = TPE)) +
  geom_line(aes(colour = subj)) +
  facet_grid(~session)

indiv_tpe <- ggplot(data, aes(x = Time, y = TPE, colour = session, linetype = session)) +
  geom_line(alpha = 0.7) +
  scale_colour_manual(values = c('deepskyblue3','springgreen3'), labels = group_labels, name = "Session") +
  scale_linetype_manual(values = c("solid", "dashed"), labels=group_labels, name = "Session") +
  xlab("Time [20s]") +
  ylab("Tactical Picture Error (TPE)") +
  facet_wrap(~subj) +
  theme_light() +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size=12),
        legend.title = element_blank())

example <- data %>%
  filter(subj == "CT38")
ggplot(example, aes(x = Time, y = TPE, colour = session, linetype = session)) +
  geom_line(alpha = 0.7, size = 0.9) +
  scale_colour_manual(values = c('deepskyblue3','springgreen3'), labels = group_labels, name = "Session") +
  scale_linetype_manual(values = c("solid", "dashed"), labels=group_labels, name = "Session") +
  xlab("Time [20s]") +
  ylab("Tactical Picture Error (TPE)") +
  theme_light() +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.text = element_text(size=12),
        legend.title = element_blank())

tiff("TPE_individual.tiff", units="in", width=10, height=7, res=300)
indiv_tpe
dev.off()  

#Mean TPE across sessions (with first 500 seconds)
mean_tpe <- aggregate(data$TPE, by = list(data$subj, data$session), FUN = mean) %>%
  dplyr::rename("subj" = Group.1,
         "session" = Group.2,
         "TPE" = x)

mean_tpe %>%
  filter(session == 1) %>%
  describe()

ggplot(mean_tpe, aes(x = session, y = TPE, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = .1) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(fill = session, group = subj), size = 3, shape = 19, alpha = 0.3, colour = "black", position = position_dodge(0.4)) +
  ylab("Tactical Picture Error") +
  ggtitle("Mean TPE Scores Across Sessions") +
  scale_x_discrete(labels=c("Pre-Meditiation", "Post-Meditation")) +
  theme_bw() + 
  theme(legend.position="none",
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.title = element_blank(),
        axis.text.y = element_text(size=12),
        axis.text.x = element_text(size=12))

#Mean TPE across sessions (without first 500 seconds)
mean_tpe500 <- filter(data, Time > 500)

mean_500 <- aggregate(mean_tpe500$TPE, by = list(mean_tpe500$subj, mean_tpe500$session), FUN = mean) %>%
  dplyr::rename("subj" = Group.1,
         "session" = Group.2,
         "TPE" = x)

means <- ggplot(mean_500, aes(x = session, y = TPE, fill = session, group = session)) +
  geom_boxplot(alpha=0.4, fill = c('turquoise2','seagreen3'), outlier.shape = NA, coef = 0, width = .3) +
  geom_line(aes(group = subj), alpha = 0.3, size = 0.4, position = position_dodge(0.4)) +
  geom_point(aes(fill = session, group = subj), size = 2, shape = 19, alpha = 0.2, colour = "black", position = position_dodge(0.4)) +
  ylab("Mean TPE") +
  ggtitle("A. Mean TPE across Session") +
  scale_fill_manual(values = c('turquoise2','seagreen3'), labels=group_labels, name = "Session") +
  scale_x_discrete(labels=group_labels, name = "Session") +
    scale_y_continuous(limits = c(0, 18000)) +
  theme_bw() + 
  theme(legend.position="none",
        plot.title = element_text(size=14, face="bold"),
        panel.border = element_rect(color="black", fill=NA, size = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size=12),
        axis.text.x = element_text(size=12))

means 

#Mean TPE across Time and Sessions, collapsed for all participants
avg_time <- data %>%
  dplyr::select(TPE,Time,session) %>%
  group_by(Time,session) %>%
  dplyr::summarise(mean_TPE = mean(TPE),
            se = (sd(TPE))/sqrt(n())) %>% 
  mutate(Time = as.numeric(Time))

over_time <- ggplot(avg_time, aes(x=Time, y=mean_TPE, fill=session, linetype = session))+
  geom_smooth(color="black", size=0.5) +
  scale_color_manual(values = c('turquoise2','seagreen3'), labels=group_labels, name = "Session") +
  scale_fill_manual(values = c('turquoise2','seagreen3'), labels=group_labels, name = "Session") +
  xlab("Time [20s]") + 
  ylab("Mean Tactical Picture Error (TPE)") +
  ggtitle("B. Mean TPE across Task Duration") +
  scale_linetype_manual(values = c("solid", "dashed"), labels=group_labels, name = "Session") +
  scale_x_continuous(limits = c(0, 2375)) +
  theme_bw() +
  theme(axis.title.x = element_text(size=14),
                  plot.title = element_text(size=14, face="bold"),
                  axis.text.x = element_text(size=12),
                  panel.border = element_rect(color="black", fill=NA, size = 1),
                  axis.text.y = element_text(size=12),
                  legend.position = c(0.7, 0.8),
                  legend.text = element_text(size = 12),
                  legend.title = element_blank())

over_time

tiff("means.tiff", units="in", width=10, height=5, res=300)
ggarrange(means, over_time, widths = c(6, 7), common.legend = FALSE)
dev.off()

#ggarrange all plots
tiff("all_session.tiff", units="in", width=12, height=8, res=300)
ggarrange(means, over_time, paf.plot, cog.plot, slope.plot, intercept.plot, nrow = 2, ncol = 3, common.legend = FALSE)
dev.off()

```

#Creating Decline Variable and Baseline information for models

```{r extract individual regression slopes between TPE and Time for first 500 seconds}

# extract the individual regression slopes for TPE and Time
# start with the practice session
regression_data1 <- data %>%
  dplyr::select(subj, TPE, Time, session) %>%
  filter (session == "1" & Time < 500) %>%
  dplyr::select(-session)

regression_data2 <- data %>%
  dplyr::select(subj, TPE, Time, session) %>%
  filter (session == "2" & Time < 500) %>%
  dplyr::select(-session)

tpe_coefs1 = c()

for (i in c(unique(regression_data1$subj))) {
  # Create temporary data frame:
  tpe_tmp <- regression_data1[regression_data1$subj==i,]
  # Perform regression:
  reg_result <- lm(tpe_tmp$TPE ~ tpe_tmp$Time)
  # Get coefficient:
  tmp_coef <- coef(reg_result)
  # Store coefficient:
  tpe_coefs1[i] <- tmp_coef[2]
}

tpe_coefs2 = c()

for (i in c(unique(regression_data2$subj))) {
  # Create temporary data frame:
  tpe_tmp <- regression_data2[regression_data2$subj==i,]
  # Perform regression:
  reg_result <- lm(tpe_tmp$TPE ~ tpe_tmp$Time)
  # Get coefficient:
  tmp_coef <- coef(reg_result)
  # Store coefficient:
  tpe_coefs2[i] <- tmp_coef[2]
}

decline1 <- data.frame(tpe_coefs1) %>%
  rownames_to_column(var = "subj") %>%
  mutate(session = "1") %>%
  dplyr::rename(decline = "tpe_coefs1")

decline2 <- data.frame(tpe_coefs2) %>%
  rownames_to_column(var = "subj") %>%
  mutate(session = "2") %>%
  dplyr::rename(decline = "tpe_coefs2")

decline <- full_join(decline1, decline2)

# now filter the original data frame to remove the first 500 seconds and join the "decline" data frame to the model
data_all <- data %>%
  filter(Time >= 500) %>%
  left_join(decline, by = c("subj", "session"))

```

```{r creating a baseline column ad exploring data}

#Create baseline TPE score
a <- data_all %>% filter(session == "1") %>%
  dplyr::rename(baseline = "TPE", in_slope = "slope", in_intercept = "intercept", in_paf = "paf", in_cog = "cog") %>%
  dplyr::select(subj, baseline, in_slope, in_intercept, in_paf, in_cog, Time, decline)

b <- data_all %>% filter(session == "2") %>%
  dplyr::rename(post_slope = "slope", post_intercept = "intercept", post_paf = "paf", post_cog = "cog") %>%
  dplyr::select(Time, subj, TPE, post_slope, post_intercept, post_paf, post_cog, training)
  
almost_baselined <- full_join(b, a, by = c("subj", "Time")) %>%
  na.omit()

baselined <- full_join(almost_baselined, maas.2, by = "subj") %>%
  filter(subj != "CT22" & subj != "CT25" & subj != "CT14" & subj != "CT39" & subj != "CT47")

baselined$subj <- as.factor(baselined$subj)

data_with_skew <- left_join(baselined, ct_skew, by = "subj")

```

```{r modelling with baseline}

#Models with just cognitive training and TPE
summary(baselined$training)
lmer_ct_withdecline <- lmer(TPE ~ ns(Time, 4) * training + baseline + decline + (1 | subj), data = baselined)

#Testing with random slopes (did not converge)
#lmer2 <- lmer(TPE ~ training + baseline + decline + (Time | subj), data = baselined)

Anova(lmer_ct_withdecline)
summary(lmer_ct_withdecline)
model_performance(lmer_ct_withdecline)
check_model(lmer_ct_withdecline)

data_ct <- ggemmeans(lmer_ct_withdecline,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]"),
                     ci.lvl = 0.83)

lmer_ct.effects <- data_ct %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")))

plot.ct <- ggplot(lmer_ct.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  ylab("Tactical Picture Error") +
  ggtitle("A. Effect of Adherence on Performance") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.ct

tiff("model_enagegement_with_baseline.tiff", units="in", width=10, height=8, res=300)
plot.ct
dev.off()

#Model with cognitive training, tpe and MAAS scores
summary(baselined$MAAS)

lmer_maas <- lmer(TPE ~ ns(Time, 4) * training * MAAS + decline + baseline + (1 | subj), data = baselined)

Anova(lmer_maas)
summary(lmer_maas)
model_performance(lmer_maas)
check_model(lmer_maas)

data_maas <- ggemmeans(lmer_maas,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "MAAS [2.83, 3.494, 4.0]"),
                     ci.lvl = 0.83)

lmer_maas.effects <- data_maas %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")),
         MAAS = factor(facet, labels = c("Low MAAS","Moderate MAAS","High MAAS")))

plot.maas <- ggplot(lmer_maas.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~MAAS) +
  ylab("Tactical Picture Error") +
  ggtitle("A. Interaction between MAAS scores and Adherence") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.maas

#Plot with geom-smooth()
ggplot(lmer_maas.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_smooth() +
  facet_grid(~MAAS) +
  ylab("Tactical Picture Error") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

tiff("model_enagegement_with_MAAS.tiff", units="in", width=12, height=8, res=300)
plot.maas
dev.off()

#Model with initial slope * training
summary(baselined$in_slope)

lmer_slope <- lmer(TPE ~ ns(Time, 4) * in_slope * training + baseline + decline + (1 | subj), data = baselined)

Anova(lmer_slope)
summary(lmer_slope)
model_performance(lmer_slope)
check_model(lmer_slope)

data_inslope <- ggemmeans(lmer_slope,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "in_slope [-1.23, -1.02, -0.83]"), 
                     ci.lvl = 0.83)

lmer_slope.effects <- data_inslope %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence"))) %>%
  mutate(in_slope = factor(facet, labels = c("Steep","Moderate","Flatter")))

plot.slope <- ggplot(lmer_slope.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~in_slope) +
  ylab("Tactical Picture Error") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  ggtitle("B. Interaction between 1/f Slope and Adherence") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.slope

tiff("model_slope_engagement.tiff", units="in", width=10, height=8, res=300)
plot.slope
dev.off()

#Model with initial IAF * training
summary(baselined$in_cog)
lmer_IAF <- lmer(TPE ~ ns(Time, 4) * in_cog * training  + baseline + decline + (1 | subj), data = baselined)

Anova(lmer_IAF)
summary(lmer_IAF)
model_performance(lmer_IAF)
check_model(lmer_IAF)

data_inIAF <- ggemmeans(lmer_IAF,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "in_cog [9.25, 9.924, 10.5]"), 
                     ci.lvl = 0.83)

lmer_IAF.effects <- data_inIAF %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence"))) %>%
  mutate(in_slope = factor(facet, labels = c("Low (<9.75)","Moderate (10.25)","High (>10.5)")))

plot.IAF <- ggplot(lmer_IAF.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~in_slope) +
  ylab("Tactical Picture Error") +
  ggtitle("B. Interaction between IAF and Adherence") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.IAF

tiff("model_iaf_engagament.tiff", units="in", width=10, height=8, res=300)
plot.IAF
dev.off()

#Initial Intercept on Performance
summary(baselined$in_intercept)

lmer_int <- lmer(TPE ~ ns(Time, 4) * in_intercept * training  + baseline + decline + (1 | subj), data = baselined)

Anova(lmer_int)
summary(lmer_int)
model_performance(lmer_int)
check_model(lmer_int)

data_inint <- ggemmeans(lmer_int,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "in_intercept [-25.79, -25.11, -24.50]"), 
                     ci.lvl = 0.83)

lmer_int.effects <- data_inint %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence"))) %>%
  mutate(in_intercept = factor(facet, labels = c("High","Moderate","Low")))

plot.int <- ggplot(lmer_int.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~in_intercept) +
  ylab("Tactical Picture Error") +
  ggtitle("D. Interaction between 1/f Intercept and Adherence") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.int

tiff("model_iaf_engagament.tiff", units="in", width=10, height=8, res=300)
plot.IAF
dev.off()

tiff("ct_maas_neural_models.tiff", units="in", width=15, height=10, res=300)
ggarrange(plot.maas, plot.IAF, plot.slope, plot.int, common.legend = TRUE, ncol = 2, nrow = 2, legend = "bottom")
dev.off()

```

##Modelling time of adherence (i.e., using skew)

```{r skew models}

#Models with cognitive training and TPE and skew (i.e., time of meditation practice)
summary(data_with_skew$skew)
describe(ct_skew)

ggplot(data_with_skew, aes(x = skew)) +
  geom_histogram()

lmer_skew <- lmer(TPE ~ ns(Time, 4) * training * skew + baseline + decline + (1 | subj), data = data_with_skew)

Anova(lmer_skew)
summary(lmer_skew)
model_performance(lmer_skew)
check_model(lmer_skew)

data_ct_skew <- ggemmeans(lmer_skew,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "skew [-1, 0, 1]"),
                     ci.lvl = 0.83)

lmer_skew.effects <- data_ct_skew %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")),
         when = factor(facet, labels = c("Negative Skew (Late Adherence)","Consistent Practice","Positive Skew (Early Adherence)")))

plot.skew <- ggplot(lmer_skew.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_wrap(~when) +
  ylab("Tactical Picture Error") +
  ggtitle("B. Interaction of Adherence and Skew") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="bottom",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 10))

plot.skew

tiff("model_enagegement_with_skew.tiff", units="in", width=10, height=8, res=300)
plot.skew
dev.off()

```

#Modelling how neural metrics change according to engagement in cognitive training
```{r models for predicting neural metric change}

#preparing dataframe (don't need all tpe metrics)
neural_metrics <- left_join(iaf_ap, ct, by = "subj") %>%
  na.omit()

neural_a <- neural_metrics %>% 
  filter(session == "1") %>%
  rename(in_slope = "slope", in_intercept = "intercept", in_paf = "paf", in_cog = "cog") %>%
  dplyr::select(subj, in_slope, in_intercept, in_paf, in_cog)

neural_b <- neural_metrics %>% filter(session == "2") %>%
  rename(post_slope = "slope", post_intercept = "intercept", post_paf = "paf", post_cog = "cog", training = "total") %>%
  dplyr::select(subj, post_slope, post_intercept, post_paf, post_cog, training)
  
neural_almost <- left_join(neural_b, neural_a, by = c("subj")) %>%
  na.omit()

neural_data <- neural_almost %>%
  mutate(IAFchange = post_cog - in_cog,
         slopechange = post_slope - in_slope,
         intchange = post_intercept - in_intercept)

describe(neural_data)

#Modelling second neural metric from cognitive training
lm.nc_IAF <- lm(post_cog ~ training + in_cog, data = neural_data)
plot(lm.nc_IAF)
summary(lm.nc_IAF)

lm.nc_int <- lm(post_intercept ~ training + in_intercept, data = neural_data)
summary(lm.nc_int)
Anova(lm.nc_int)

lm.nc_slope <- lm(post_slope ~ in_slope + training, data = neural_data)
summary(lm.nc_slope)
Anova(lm.nc_slope)

```

#Additional Analyses following April Meeting

```{r personality}
#Assessing whether enjoyment plays a role in benefits gained
enjoy <- read_rds("enjoyment_summary.rds")

range <- enjoy %>%
  group_by(subj) %>%
  count() %>%
  mutate(observations = n/5)
  
enjoy_summary <- enjoy %>%
  group_by(subj, measure) %>%
  summarise(means = mean(score)) %>%
  filter(subj != "CT22" & subj != "CT25" & subj != "" & subj != "CT52")

ggplot(enjoy_summary, aes(x = measure, y = means)) +
  geom_boxplot(aes(fill = measure), alpha = 0.3)

to_join_enjoy <- pivot_wider(enjoy_summary, names_from = "measure", values_from = means)
  
baselined_enjoy <- left_join(baselined, to_join_enjoy, by = "subj")

summary(baselined_enjoy$enjoy)
lmer_enjoy_maas <- lmer(TPE ~ ns(Time, 4) * training * enjoy + baseline + decline + (1|subj), data = baselined_enjoy)
summary(lmer_enjoy_maas)
Anova(lmer_enjoy_maas)

data_enjoym <- ggemmeans(lmer_enjoy_maas,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "enjoy [58.33, 71.42, 85.57]"),
                     ci.lvl = 0.83)

lmer_enjoym.effects <- data_enjoym %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")),
         enjoy = factor(facet, labels = c("Low Enjoyment", "Moderate Enjoyment", "High Enjoyment")))

plot.enjoym <- ggplot(lmer_enjoym.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~enjoy) +
  ylab("Tactical Picture Error") +
  xlab("Time [20s]") +
  ggtitle("D. Interaction between Adherence and Enjoyment") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.enjoym

tiff("model_enagegement_with_enjoy.tiff", units="in", width=10, height=8, res=300)
plot.enjoym
dev.off()

#Personality Effects
personality_sum <- read_rds('personality_summary.rds')

pn_TPE_maas <- personality_sum %>%
  left_join(neural_metrics, by = "subj") %>%
  left_join(mean_500, by = c("subj", "session")) %>%
  left_join(maas.2, by = "subj")

pn_enjoy <- left_join(pn_TPE_maas, to_join_enjoy, by = "subj")

#Check how conscientiousness is related to adherence (relevant to reviewer comment)
tiff("conscxengagement.tiff", units="in", width=6, height=6, res=300)
pn_TPE_maas %>% 
  filter(session == "1") %>%
  na.omit() %>%
  ggplot(aes(x = total, y = Conscientiousness)) +
  geom_jitter(shape=20, size = 3, alpha = 0.5) +
  geom_smooth(method=lm, fill = "lightgreen", colour = "black", linetype = "dotted") +
  labs(y="Conscientiousness Score", x = "Cognitive Training Engagement") +
  theme_light() +
  theme(axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.text.x = element_text(size=12),
        plot.title = element_text(size=14, face="bold"),
        axis.text.y = element_text(size=12)) +
  stat_regline_equation() +
  stat_cor(label.y = 38)
dev.off()

pn_enjoy_sk <- select(ct_skew, skew, subj) %>%
  left_join(pn_enjoy, by = "subj")

cplot_training <- pn_enjoy_sk %>% 
  filter(session == "1") %>%
  dplyr::select("Conscientiousness", "Neuroticism", "enjoy", "skew" , "enjoy", "MAAS", "total", "TPE", "slope", "cog", "intercept") %>%
  rename(Slope = slope, IAF = cog, Intercept = intercept, Enjoy = enjoy, Skew = skew, Adherence = total)

cplot_enjoy <- pn_enjoy %>% 
  filter(session == "2") %>%
  dplyr::select("calm", "distract", "energy", "enjoy", "success" , "Conscientiousness", "Neuroticism", "Mind-Wandering", "MAAS", "slope", "cog", "total", "TPE")

#Looking at all correlations between variables
library(corrplot)
library(correlation)

cor_data = cor(cplot_training, use="pairwise.complete.obs")

tiff("corrplot.tiff", units="in", width=12, height=7, res=300)
corrplot(cor_data, method="color", col=colorRampPalette(c("lightskyblue3","white","springgreen3"))(8),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
dev.off()

pairs.panels(cplot_training,
             smooth = TRUE,      # If TRUE, draws loess smooths
             scale = FALSE,      # If TRUE, scales the correlation text font
             ellipses = FALSE,
             density = TRUE,     # If TRUE, adds density plots and histograms
             method = "pearson", # Correlation method (also "spearman" or "kendall")
             pch = 21,           # pch symbol
             lm = TRUE,         # If TRUE, plots linear fit rather than the LOESS (smoothed) fit
             cor = TRUE,         # If TRUE, reports correlations
             jiggle = FALSE,     # If TRUE, data points are jittered
             factor = 2,         # Jittering factor
             hist.col = 2,       # Histograms color
             stars = TRUE,       # If TRUE, adds significance level with stars
             ci = TRUE) 

pairs.panels(cplot_enjoy,
             smooth = TRUE,      # If TRUE, draws loess smooths
             scale = FALSE,      # If TRUE, scales the correlation text font
             ellipses = FALSE,
             density = TRUE,     # If TRUE, adds density plots and histograms
             method = "pearson", # Correlation method (also "spearman" or "kendall")
             pch = 21,           # pch symbol
             lm = TRUE,         # If TRUE, plots linear fit rather than the LOESS (smoothed) fit
             cor = TRUE,         # If TRUE, reports correlations
             jiggle = FALSE,     # If TRUE, data points are jittered
             factor = 2,         # Jittering factor
             hist.col = 2,       # Histograms color
             stars = TRUE,       # If TRUE, adds significance level with stars
             ci = TRUE) 

person_model <- pn_TPE_maas %>%
  janitor::clean_names() 

baselined_person <- person_model %>%
  filter(session == "1") %>%
  dplyr::select(conscientiousness, neuroticism, openness_to_experience, subj) %>%
  left_join(baselined, by = "subj")

describe(baselined_person)

#Modelling two personality variables
lmer_big2 <- lmer(TPE ~ ns(Time, 4) * training * neuroticism * conscientiousness + baseline + decline + (1|subj), data = baselined_person)

summary(lmer_big2)
Anova(lmer_big2)
model_performance(lmer_big2)
check_model(lmer_big2)

summary(baselined_person$neuroticism)
summary(baselined_person$conscientiousness)
summary(baselined_person$MAAS)

data_ct_big5 <- ggemmeans(lmer_big2,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "conscientiousness [27, 29.86, 33]", "neuroticism [16, 20, 19.7]"),
                     ci.lvl = 0.83)

lmer_big5.effects <- data_ct_big5 %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")),
         conscientiousness = factor(facet, labels = c("Low Consc","Moderate Consc","High Consc")),
         neuro = factor(panel, labels = c("Low Neur", "Moderate Neur", "High Neur")))

plot.big5 <- ggplot(lmer_big5.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(conscientiousness ~ neuro) +
  ylab("Tactical Picture Error") +
  ggtitle("Interaction of Personality Factors and Adherence") +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000)) +
  xlab("Time [20s]") +
  theme_bw() +
  theme(legend.position="bottom",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.big5

#assessing if prior to sleep if useful
sleep <- read.csv("sleep-mindful.csv") %>%
  mutate(prior = as.factor(prior))

summary(sleep)

baselined_sleep <- left_join(baselined, sleep, by = "subj")

#Modelling prior sleep and MAAS
lmer_sleep <- lmer(TPE ~ ns(Time, 4) * training * prior + baseline + decline + (1|subj), data = baselined_sleep)
summary(lmer_sleep)
Anova(lmer_sleep)

data_sleepm <- ggemmeans(lmer_sleep,
                     terms = c("Time [all]", "training [4.00, 10.28, 16.00]", "prior[all]"),
                     ci.lvl = 0.83)

lmer_sleepm.effects <- data_sleepm %>% 
  as_tibble() %>% 
  mutate(training = factor(group, labels = c("Low Adherence","Moderate Adherence","Strict Adherence")),
         sleep = factor(facet, labels = c("Prior to First Sleep", "After")))

plot.sleep <- ggplot(lmer_sleepm.effects, aes(x,predicted, ymin = conf.low, ymax = conf.high)) +
  geom_line(aes(linetype=training)) +
  scale_linetype_manual(values=c("solid", "longdash", "dotdash"), name = NULL) +
  scale_fill_manual(values = c('lightskyblue1','palegreen',"tan"), name = NULL) +
  scale_colour_manual(values = c('slateblue','springgreen3',"violetred")) +
  geom_ribbon(aes(fill=training),alpha=0.3) +
  facet_grid(~sleep) +
  ylab("Tactical Picture Error") +
  xlab("Time [20s]") +
  ggtitle("C. Interaction of Sleep and Adherence") +
  theme_bw() +
  theme(legend.position="top",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=16, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 14))

plot.sleep

#Assessing model fits 
table <- compare_performance(lmer_big2, lmer_sleep, lmer_ct_withdecline, lmer_IAF, lmer_skew, lmer_int, lmer_maas, rank = TRUE)

print(table)
table <- compare_performance(lmer_consc_neuroticism, lmer_consc_neurotcism_maas, lmer_neuroticism_maas, lmer_ct_withdecline, lmer_maas_iaf, lmer_ct_withdecline, lmer_enjoy_maas, lmer_sleep_maas, lmer_skew, lmer_conscientiousness_maas, lmer_maas, lmer_IAF, lmer_slope, lmer_int,  rank = TRUE)

summary(tmod)$table

#Creating Plots for Manuscript
tiff("ct_individual_factors.tiff", units="in", width=15, height=10, res=300)
ggarrange(plot.ct, plot.skew, plot.sleep, plot.enjoym, common.legend = TRUE, ncol = 2, nrow = 2, legend = "bottom")
dev.off()

tiff("big5.tiff", units="in", width=9, height=8, res=300)
plot.big5
dev.off()
```

#Create Model Output Tables for Latex

```{r create tables}

library(lmerOut)
library(kableExtra)

#models with just cognitive training and TPE
lmer_ct_withdecline <- lmer(TPE ~ ns(Time, 4) * training + baseline + decline + (1 | subj), data = baselined)
kable(Anova(lmer_ct_withdecline), "latex", caption = "Summary of effects of Engagement on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/engagement.pdf", keep_tex = TRUE)
pprint(lmer_ct_withdecline, "data-csvs/tables/engagement_summary.tex")

#model with cognitive training, tpe and MAAS scores
lmer_maas <- lmer(TPE ~ ns(Time, 4) * training * MAAS + decline + baseline + (1 | subj), data = baselined)
kable(Anova(lmer_maas), "latex", caption = "Summary of effects of Engagement and MAAS on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/maas.pdf", keep_tex = TRUE)
pprint(lmer_maas, "data-csvs/tables/maas_summary.tex")

#Model with initial slope * training
lmer_slope <- lmer(TPE ~ ns(Time, 4) * in_slope * training + baseline + decline + (1 | subj), data = baselined)
kable(Anova(lmer_slope), "latex", caption = "Summary of effects of Engagement and 1/f Slope on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/slope.pdf", keep_tex = TRUE)
pprint(lmer_slope, "data-csvs/tables/slope_summary.tex")

#Model with initial IAF * training
lmer_IAF <- lmer(TPE ~ ns(Time, 4) * in_cog * training  + baseline + decline + (1 | subj), data = baselined)
kable(Anova(lmer_IAF), "latex", caption = "Summary of effects of Engagement and IAF on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/iaf.pdf", keep_tex = TRUE)
pprint(lmer_IAF, "data-csvs/tables/iaf_summary.tex")

#Initial Intercept on Performance
lmer_int <- lmer(TPE ~ ns(Time, 4) * in_intercept * training  + baseline + decline + (1 | subj), data = baselined)
kable(Anova(lmer_int), "latex", caption = "Summary of effects of Engagement and 1/f Intercept on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/intercept.pdf", keep_tex = TRUE)
pprint(lmer_int, "data-csvs/tables/intercept_summary.tex")

#Model with Skew
lmer_skew <- lmer(TPE ~ ns(Time, 4) * training * skew + baseline + decline + (1 | subj), data = data_with_skew)
kable(Anova(lmer_skew), "latex", caption = "Summary of effects of Engagement and Skew on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/skew.pdf", keep_tex = TRUE)
pprint(lmer_skew, "data-csvs/tables/skew_summary.tex")

#Enjoyment Model
lmer_enjoy_maas <- lmer(TPE ~ ns(Time, 4) * training * enjoy + baseline + decline + (1|subj), data = baselined_enjoy)
kable(Anova(lmer_enjoy_maas), "latex", caption = "Summary of effects of Engagement and Enjoyment on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/enjoy.pdf", keep_tex = TRUE)
pprint(lmer_enjoy_maas, "data-csvs/tables/enjoyment_summary.tex")

#Personality model
lmer_big2 <- lmer(TPE ~ ns(Time, 4) * training * neuroticism * conscientiousness + baseline + decline + (1|subj), data = baselined_person)
kable(Anova(lmer_big2), "latex", caption = "Summary of effects of Engagement, Neuroticism and Conscientiousness on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/big2.pdf", keep_tex = TRUE)
pprint(lmer_big2, "data-csvs/tables/big2_summary.tex")

#prior sleep and MAAS
lmer_sleep <- lmer(TPE ~ ns(Time, 4) * training * prior + baseline + decline + (1|subj), data = baselined_sleep)
kable(Anova(lmer_sleep), "latex", caption = "Summary of effects of Engagement and Sleep on TPE Scores (Type II Wald Tests)", booktabs = T) %>%
  save_kable("data-csvs/tables/sleep.pdf", keep_tex = TRUE)
pprint(lmer_sleep, "data-csvs/tables/sleep_summary.tex")

```
